---
title: "On the origin of citations"
author: "Ali Yesilkanal"
date: "Tuesday, September 15, 2015"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: espresso
    fig_width: 8
    fig_height: 8
    fig_caption: yes
    self_contained: false
---

#Load the data

```{r load_data}
counts_raw <- read.delim("data/counts-raw.txt.gz")
counts_norm <- read.delim("data/counts-norm.txt.gz")
```

#Data Exploration

What is the distribution of authors in all articles of our data set?

```{r author_histogram, fig.cap= "Figure 1: Number of Authors per paper", echo=FALSE}
hist(counts_raw$authorsCount, main="Authors per paper", xlab="# authors")
```

What is the number of facebook shares per article?

```{r fbshare_histogram, fig.cap= "Figure 2: Number of Facebook shares per article", echo=FALSE}
hist(counts_raw$facebookShareCount, main="Facebook shares per paper", xlab="# Facebook shares")
```

The average number of Facebook shares per paper in the data set is `r mean(counts_raw$facebookShareCount)`.

##dplyr

```{r}
library("dplyr")
```

```{r}
research <- filter(counts_raw, articleType == "Research Article")
```

```{r}
research_2006 <- filter(research, year == 2006)
nrow(research_2006)
```

```{r}
research_2006_fb <- filter(research, year==2006, facebookCommentCount > 0)
nrow(research_2006_fb)
```

```{r}
research_2006_fb_tweet_disease <- filter(research, year==2006, 
                           facebookCommentCount > 0 | backtweetsCount > 0,
                           grepl("Infectious Diseases", plosSubjectTags))
nrow(research_2006_fb_tweet_disease)
```


```{r}
colnames(research)
```

```{r}
article_info <- select(research, doi, pubDate, journal, title, articleType, authorsCount)
article_info <- select(research, doi:authorsCount)
colnames(article_info)
```

```{r}
metrics <- select(research, contains("Count"), -authorsCount, f1000Factor, wikipediaCites)
colnames(metrics)
```

```{r}
head(select(research, journal))
head(select(research, 3))
```

```{r}
slice(article_info, 1:3)
```

```{r challenge2}
low_cite <- filter(research, year <= 2008, pdfDownloadsCount > 1000, mendeleyReadersCount > 15,
                   wosCountThru2011 < 10)
nrow(low_cite)
data <- select(low_cite, journal, title, year)
data
```

### Chaining commands with dplyr (piping data)

pipe character in R is "%>%"

```{r}
facebook_2006 <- research %>% filter(year == 2006) %>% 
  select(contains("facebook"))
head(facebook_2006)
```

```{r}
research %>% filter(year == 2006) %>% 
  select(contains("facebook")) %>%
  head
```

```{r}
research %>% arrange(authorsCount, wosCountThru2011) %>%
  select(authorsCount,wosCountThru2011) %>%
  slice(1:10)
```

```{r}
research %>% arrange(desc(authorsCount), desc(wosCountThru2011)) %>%
  select(authorsCount,wosCountThru2011) %>%
  slice(1:10)
```

```{r challenge13.1}
research %>% select(wosCountThru2011, title) %>% arrange(desc(wosCountThru2011)) %>% head
```

```{r challenge13.2}
research %>% select(authorsCount, title, journal, plosSubjectTags) %>% arrange(desc(authorsCount)) %>% head
```

### Summarizing with dplyr

```{r}
research <- research %>% mutate(weeksSincePublished = daysSincePublished / 7,
                                yearsSincePublished = weeksSincePublished / 52)
research %>% select(contains("Since")) %>% slice(1:10)
```

using summarize

```{r}
research %>% summarize(plos_mean = mean(plosCommentCount),
                       plos_sd = sd(plosCommentCount),
                       num = n())
```


### Using group_by 

```{r}
research %>% group_by(journal) %>% summarize(tweets_mean = mean(backtweetsCount))
```

```{r}
research %>% group_by(journal, year) %>% summarize(tweets_mean = mean(backtweetsCount))
```

```{r challenge14}
tweets_per_journal <- research %>% group_by(journal) %>% 
  summarize(num = n(), 
            mean = mean(backtweetsCount),
            sem = sd(backtweetsCount) / sqrt(num))
tweets_per_journal                                                          
```


# ggplot2

```{r}
library("ggplot2")
```

```{r}
p <- ggplot(research, aes(x = pdfDownloadsCount,
                                    y = wosCountThru2011)) +
  geom_point(aes(color = journal))
p
```

```{r}
p <- ggplot(research, aes(x = pdfDownloadsCount,
                                    y = wosCountThru2011)) +
  geom_point(aes(size = authorsCount))
p
```

```{r}
p <- ggplot(research, aes(x = pdfDownloadsCount,
                                    y = wosCountThru2011)) +
  geom_point(aes(alpha = daysSincePublished))
p
```
aes() is required only if you want the plot aesthetics to be dependent on some property of the data. 

```{r}
p <- ggplot(research, aes(x = pdfDownloadsCount,
                                    y = wosCountThru2011)) +
  geom_point(aes(color = journal)) +
  geom_smooth()
p
```

```{r}
p <- ggplot(research, aes(x = pdfDownloadsCount,
                          y = wosCountThru2011,
                          color = journal)) +
  geom_point() +
  geom_smooth()
  
p
```

```{r challenge15}
p <- ggplot(research, aes(x = daysSincePublished,
                          y = wosCountThru2011,)) +
  geom_point(aes(color = journal), alpha=0.5) +
  geom_smooth(color="red")
  
p

```

### Using scales

```{r}
p <- ggplot(research, aes(x = pdfDownloadsCount,
                          y = wosCountThru2011,
                          )) +
  geom_point(aes(color = journal)) +
  geom_smooth()  
p + scale_x_log10() + scale_y_log10()
```
cannot use the sooth line with this one. to counteract that, take the logs earlier

```{r}
p <- ggplot(research, aes(x = log10(pdfDownloadsCount +1),
                          y = log10(wosCountThru2011 + 1),
                          )) +
  geom_point(aes(color = journal)) +
  geom_smooth()  
p 
```

```{r}
p <- ggplot(research, aes(x = log10(pdfDownloadsCount +1),
                          y = log10(wosCountThru2011 + 1),
                          )) +
  geom_point(aes(color = journal)) +
  geom_smooth() +
  scale_x_continuous(breaks = c(1,3), labels = c(10,1000)) +
  scale_y_continuous(breaks = c(1,3), labels = c(10,1000))
p 
```

different color options

```{r}
p + scale_colour_grey()
p + scale_color_manual(values = c("red", "green","yellow", "orange","purple", "blue", "pink"))
```

```{r}
library("RColorBrewer")
display.brewer.all(type = "qual")
p + scale_color_brewer(palette ="Dark2")
```

```{r}
p + scale_color_brewer(palette ="Dark2",
                       labels = 1:7, name = "PLOS")
```

```{r challenge16}
p <- ggplot(research, aes(x = sqrt(pdfDownloadsCount),
                          y = sqrt(wosCountThru2011),
                          )) +
  geom_point(aes(color = journal)) +
  geom_smooth() +
  scale_color_brewer(palette = "Accent")
p
```

### Using facets to make subplots

```{r}
p <- ggplot(research, aes(x = sqrt(pdfDownloadsCount),
                          y = sqrt(wosCountThru2011),
                          )) +
  geom_point(aes(color = journal)) +
  geom_smooth() +
  scale_color_brewer(palette = "Accent")
p + facet_wrap(~journal, ncol = 2 )
```

## Using facet_grid

```{r}
research <- mutate(research, immuno = grepl("Immunology", plosSubjectTags))
p <- ggplot(research, aes(x = sqrt(pdfDownloadsCount),
                          y = sqrt(wosCountThru2011),
                          )) +
  geom_point(aes(color = journal)) +
  geom_smooth() +
  scale_color_brewer(palette = "Accent") +
  facet_grid(journal~immuno)
p
```

## Using different geoms

### boxplot
```{r}
p <- ggplot(research, aes(x = journal,
                          y = sqrt(wosCountThru2011),
                          )) +
  geom_boxplot()
p
```


### barplot
```{r}
tweets_per_journal <- research %>% group_by(journal) %>% 
  summarize(num = n(), 
            mean = mean(backtweetsCount),
            sem = sd(backtweetsCount) / sqrt(num))

tweet_bar <- ggplot(tweets_per_journal, aes(x=journal, y=mean)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin= mean - sem, ymax= mean + sem), width = 0.1) +
  geom_text(aes(label = num), hjust = 0, vjust= 0)
tweet_bar

```

```{r challenge18}
tweets_per_journal_year <- research %>% group_by(journal, year) %>% 
  summarize(num = n(), 
            mean = mean(backtweetsCount),
            sem = sd(backtweetsCount) / sqrt(num))
tweets_per_journal_year

tweet_bar <- ggplot(tweets_per_journal_year, aes(x=journal, y=mean)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin= mean - sem, ymax= mean + sem), width = 0.1) +
  geom_text(aes(label = num)) +
  facet_wrap(~year)
tweet_bar
```

###  Customizing the plot

```{r}
tweet_bar + labs(title="mean tweets per journal per year", x="journal", y="mean number of tweets") +
  theme_bw()
```

```{r}
theme_set(theme_bw())
```

